European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-12 manuscript/manuscript.qmd
Better keywords and a draft for conclusion (#62)
* Initial plan

* Improved keywords and added comprehensive conclusion section

Co-authored-by: kv9898 <105025148+kv9898@users.noreply.github.com>

* update keywords

* fix 2nd para

* Major: fix 3rd para to avoid misinterpretation

* Major: fix 9th para to soften the tone

* fix 10th para

* remove the para on 3 european powers

* Major: fix 3rd limitation

* future research: 1st

* future research: 2nd

* future research: 3rd

* Fix closing

In conclusion, this paper demonstrates that the International Monetary Fund, despite its technocratic veneer, remains deeply shaped by geopolitical considerations. However, the nature of great power influence is more complex and differentiated than often assumed. European powers exercise broad institutional influence over access to IMF resources, while American power manifests more selectively through conditionality. This division of influence has important implications for understanding both the current operations of the international financial system and the prospects for its future reform. As emerging economies continue to challenge Western dominance in global governance, recognising the internal dynamics and potential fractures within the Western coalition becomes increasingly important for scholars, policymakers, and reform advocates alike.

>>>>

To summarise, our analysis underscores that Western influence within the IMF is not monolithic but internally differentiated, with Europe shaping access to resources and the United States shaping the terms of their use. Recognising this asymmetry complicates prevailing narratives of IMF governance and highlights the need to take transatlantic politics seriously in studies of global financial institutions. More broadly, the paper demonstrates the value of disaggregating great power influence rather than treating blocs as unified actors. Doing so not only clarifies the institutional dynamics of the IMF, but also provides a framework for examining other organisations where overlapping alliances mask divergent interests. In this way, the findings contribute to a more nuanced understanding of how power operates in international economic governance and point toward the importance of looking inside, as well as outside, the Western coalition when considering the future of multilateralism.
